### 2019-10-07 proposal/figures/flow.tex

\documentclass{standalone}

\usepackage{tikz}
\usepackage{pgfplots}
\usepackage{tkz-euclide}

\pgfplotsset{compat=1.6}

\usetikzlibrary{positioning,calc}
\usetikzlibrary{arrows} 
\usetikzlibrary{patterns} 
\usetikzlibrary{shapes}
\usetikzlibrary{decorations.pathmorphing}

\usetkzobj{all} 
\usetikzlibrary{through}

\pgfdeclarepatternformonly{mydots}{\pgfqpoint{-1pt}{-1pt}}{\pgfqpoint{1pt}{1pt}}{\pgfqpoint{2pt}{2pt}}% original definition: \pgfqpoint{3pt}{3pt}
{%
  \pgfpathcircle{\pgfqpoint{0pt}{0pt}}{.5pt}%
  \pgfusepath{fill}%
}%

\begin{document}

\def\BeforeLight{2}                  % Light path before particle cloud
\begin{tikzpicture}

  \draw[step=0.5cm, gray!20!white, very thin] (-0.9, -0.9) grid (7.9, 3.9);

  \fill[blue!10!white, pattern=dots, pattern color=blue!20!white] (0,0) rectangle (7,3);
  \draw[line width=0.25mm] (0,0) -- (7,0);
  \draw[line width=0.25mm] (0,3) -- (7,3);

  \fill[pattern=north west lines, pattern color=blue!40!white] (1.0, 1.0) rectangle (2.0, 2.0);
  \draw[line width=0.15mm] (1.0, 1.0) rectangle (2.0, 2.0);

  \fill[pattern=north west lines, pattern color=gray] (0, 0) rectangle (7, -0.5);
  \fill[pattern=north west lines, pattern color=gray] (0, 3) rectangle (7, 3.5);

  \draw[red,
  decoration = {markings,
    mark = at position 0.18 with {\arrow[]{latex }},
    mark = at position 0.77 with {\arrow[]{latex }}}, postaction = {decorate}]
  (-1, 0.25) --++ (5.5, 0) node[midway, above, black]{};

  \draw[red,
  decoration = {markings,
    mark = at position 0.18 with {\arrow[]{latex }},
    mark = at position 0.77 with {\arrow[]{latex }}}, postaction = {decorate}]
  (-1, 0.75) --++ (5.5, 0) node[midway, above, black]{};

  \draw[red,
  decoration = {markings, width=0.3mm,
    mark = at position 0.5 with {\arrow[]{latex}}}, postaction = {decorate}]
  (-1, 1.25) --++ (2.0, 0) node[midway, above, black]{$\Gamma_{in}$};

  \draw[red,
  decoration = {markings, width=0.3mm,
    mark = at position 0.5 with {\arrow[]{latex}}}, postaction = {decorate}]
  (-1, 1.75) --++ (2.0, 0) node[midway, above, black]{};

  \draw[red,
  decoration = {markings,
    mark = at position 0.18 with {\arrow[]{latex }},
    mark = at position 0.77 with {\arrow[]{latex }}}, postaction = {decorate}]
  (-1, 2.25) --++ (5.5, 0) node[midway, above, black]{};

  \draw[red,
  decoration = {markings,
    mark = at position 0.18 with {\arrow[]{latex }},
    mark = at position 0.77 with {\arrow[]{latex }}}, postaction = {decorate}]
  (-1, 2.75) --++ (5.5, 0) node[midway, above, black]{};

  \draw[blue,
  decoration = {markings, width=0.3mm,
    mark = at position 0.5 with {\arrow[]{latex}}}, postaction = {decorate}]
  (4.5, 0.25) --++ (3.5, 0) node[midway, above, black]{};

  \draw[blue,
  decoration = {markings, width=0.3mm,
    mark = at position 0.5 with {\arrow[]{latex}}}, postaction = {decorate}]
  (4.5, 0.75) --++ (3.5, 0) node[midway, above, black]{};

  \draw[blue,
  decoration = {markings, width=0.3mm,
    mark = at position 0.5 with {\arrow[]{latex}}}, postaction = {decorate}]
  (4.5, 1.25) --++ (3.5, 0) node[midway, above, black]{$\Gamma_{out}$};

  \draw[blue,
  decoration = {markings, width=0.3mm,
    mark = at position 0.5 with {\arrow[]{latex}}}, postaction = {decorate}]
  (4.5, 1.75) --++ (3.5, 0) node[midway, above, black]{};

  \draw[blue,
  decoration = {markings, width=0.3mm,
    mark = at position 0.5 with {\arrow[]{latex}}}, postaction = {decorate}]
  (4.5, 2.25) --++ (3.5, 0) node[midway, above, black]{};

  \draw[blue,
  decoration = {markings, width=0.3mm,
    mark = at position 0.5 with {\arrow[]{latex}}}, postaction = {decorate}]
  (4.5, 2.75) --++ (3.5, 0) node[midway, above, black]{};

  \draw[red,
  decoration = {markings, width=0.3mm,
    mark = at position 0.5 with {\arrow[]{latex}}}, postaction = {decorate}]
  (2.0, 1.25) --++ (2.5, 0) node[midway, above, black]{};

  \draw[red,
  decoration = {markings, width=0.3mm,
    mark = at position 0.5 with {\arrow[]{latex}}}, postaction = {decorate}]
  (2.0, 1.75) --++ (2.5, 0) node[midway, above, black]{};

  % \draw[black!60!white, width=0.15mm] (0, -1) -- (0,4) node[at start, right, black]{$(1)$ Inflow};
  % \draw[black!60!white, width=0.15mm] (7, -1) -- (7,4) node[at start, right, black]{$(2)$ Outflow};

  % \node (A) at (4.1, -1) {$(3)$ No-slip};

  % \draw[->, line width = 0.155mm] (1.75, 1.0) -- (A)
  % \draw[->, line width = 0.155mm] (3.5, 0) -- (A)
  % \draw[->, line width = 0.155mm] (4.2, 3) -- (A)

\end{tikzpicture}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End: