### 2019-12-06 progress_update/figs/number_quest/fig.tex
Adding the number question

\documentclass[border=5mm]{standalone}
\usepackage{tikz}
\usetikzlibrary{shadows}
\usetikzlibrary{shapes}

\begin{document}
\tikzset{
  cascaded/.style = {
    general shadow = {
      shadow scale = 1,
      shadow xshift = -2ex,
      shadow yshift = 2ex,
      draw,
      thick,
      fill = white},
    general shadow = {
      shadow scale = 1,
      shadow xshift = -1ex,
      shadow yshift = 1ex,
      draw,
      thick,
      fill = white},
    fill = white, 
    draw,
    thick,
    minimum width = 2.5cm,
    minimum height = 1.3cm},
  roundnode/.style={circle, draw=green!60, fill=green!5, very thick, minimum width = 0.6cm, minimum height = 0.6cm}
}
\begin{tikzpicture}

  \node[cascaded,label=below:Time $t$] (time_t) at (0,0) {};

  \node [cloud, draw,cloud puffs=10,cloud puff arc=120, aspect=2, inner ysep=1em] (net) at (4,0) {Network};

  \node[cascaded,label=below:Time $t+1$] (time_t_1) at (8,0) {};

  \draw [->,line width=1pt, shorten >=3pt, shorten <=9pt] (time_t) edge (net);
  \draw [->,line width=1pt, shorten >=10pt, shorten <=3pt] (net) edge (time_t_1);

  \node[text width=1cm,align=center] (x_y) at (time_t.center) {Input\\ channels};
  \node[text width=1cm,align=center] (x_y_1) at (time_t_1.center) {Otput\\ channels};

  \node[roundnode] (number) at (0, -2.3) {$3.14$};

  \node[draw, circle] (quest) at (2, -0.6) {\textbf{\large ?}};
  \draw [->,line width=1pt, shorten >=3pt, shorten <=4pt] (number.east) -- (2,-2.3) -- (2,-1);

\end{tikzpicture}
\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End: