### 2019-12-09 progress_update/figs/recursive/fig.tex

\documentclass[border={7pt 0pt 0pt 4pt}]{standalone}
\usepackage{tikz}
\usetikzlibrary{shadows}
\usetikzlibrary{shapes}

\begin{document}
\tikzset{
  cascaded/.style = {
    general shadow = {
      shadow scale = 1,
      shadow xshift = -2ex,
      shadow yshift = 2ex,
      draw,
      thick,
      fill = white},
    general shadow = {
      shadow scale = 1,
      shadow xshift = -1ex,
      shadow yshift = 1ex,
      draw,
      thick,
      fill = white},
    fill = white, 
    draw,
    thick,
    minimum width = 2.5cm,
    minimum height = 1.3cm},
  roundnode/.style={circle, draw=green!60, fill=green!5, very thick, minimum width = 0.6cm, minimum height = 0.6cm}
}
\begin{tikzpicture}

  \draw [dashed, draw=blue!80, fill=blue!4, opacity=1] (-2, 1.5) rectangle (1.7, -1.3);

  \node[cascaded,label=below:Time $t$] (time_t) at (0,0) {};
  \node [cloud, draw,cloud puffs=10,cloud puff arc=120, aspect=2, inner ysep=1em] (net) at (4,0) {Network};
  \node[cascaded,label=below:Time $t+1$] (time_t_1) at (8,0) {};
  \draw [->,line width=1pt, shorten >=3pt, shorten <=9pt] (time_t) edge (net);
  \draw [->,line width=1pt, shorten >=10pt, shorten <=3pt] (net) edge (time_t_1);

  \node [cloud, draw,cloud puffs=10,cloud puff arc=120, aspect=2, inner ysep=1em] (net_1) at (4,-3) {Network};
  \draw [->,line width=1pt, shorten >=10pt, shorten <=3pt] (time_t_1.south) -- (8, -1.5) -- (0, -1.5) -- (0, -3) -- (net_1.west);
  \node[cascaded,label=below:Time $t+2$] (time_1_t_1) at (8,-3) {};
  \draw [->,line width=1pt, shorten >=10pt, shorten <=3pt] (net_1) edge (time_1_t_1);

  \node [cloud, draw,cloud puffs=10,cloud puff arc=120, aspect=2, inner ysep=1em] (net_2) at (4,-6) {Network};
  \draw [->,line width=1pt, shorten >=10pt, shorten <=3pt] (time_1_t_1.south) -- (8, -4.5) -- (0, -4.5) -- (0, -6) -- (net_2.west);
  \node[cascaded,label=below:Time $t+3$] (time_2_t_1) at (8,-6) {};
  \draw [->,line width=1pt, shorten >=10pt, shorten <=3pt] (net_2) edge (time_2_t_1);

\end{tikzpicture}
\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End: