### 2020-06-02 final/figures/network/meth.tex
Adding some nice figures

\documentclass{standalone}

\usepackage{tikz}
\usepackage{pgfplots}
\usepackage{tkz-euclide}

\pgfplotsset{compat=1.6}

\usetikzlibrary{positioning,calc}
\usetikzlibrary{arrows} 
\usetikzlibrary{patterns} 
\usetikzlibrary{shapes}
\usetikzlibrary{decorations.pathmorphing}

\begin{document}

\tikzset{
  cascaded/.style = {
    general shadow = {
      shadow scale = 1,
      shadow xshift = -2ex,
      shadow yshift = 2ex,
      draw,
      thick,
      fill = white},
    general shadow = {
      shadow scale = 1,
      shadow xshift = -1ex,
      shadow yshift = 1ex,
      draw,
      thick,
      fill = white},
    fill = white, 
    draw,
    thick,
    minimum width = 2.5cm,
    minimum height = 1.3cm},
  squarednode/.style={rectangle, draw=green!60, fill=green!5, very thick, minimum width = 2.5cm, minimum height = 1.3cm}
}

\def\BeforeLight{2}
\begin{tikzpicture}

  \node[draw,minimum width = 2.5cm, minimum height = 1.3cm, thick] at (0,3) (img) {\shortstack{Initial\\Image}};
  \node [cloud, draw,cloud puffs=10,cloud puff arc=120, aspect=2, inner ysep=1em, thick] (net) at (0, 0) {Network};
  \node[cascaded] at (0.3, -3) (images) {Images};

  \node[draw=green!60, fill=green!5, very thick, minimum width = 0.6cm, minimum height = 0.6cm] at (1.7, 3.5) (params_1) {\shortstack{Simulation\\Parameters}};

  \draw[->, shorten >=3pt, shorten <=3pt, thick ] (img) edge (net);
  \draw[->, shorten >=3pt, shorten <=3pt, thick ] (net) edge (0, -2);
  \draw[->, shorten >=3pt, shorten <=3pt, thick] (0, -1.5) -- (2, -1.5) -- (2, 2) -- (0, 2);

\end{tikzpicture}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End: