### 2020-06-02 final/figures/pipeline/meth.tex

\documentclass{standalone}

\usepackage{tikz}
\usepackage{pgfplots}
\usepackage{tkz-euclide}

\pgfplotsset{compat=1.6}

\usetikzlibrary{positioning,calc}
\usetikzlibrary{arrows} 
\usetikzlibrary{patterns} 
\usetikzlibrary{shapes}
\usetikzlibrary{decorations.pathmorphing}

\begin{document}

\tikzset{
  cascaded/.style = {
    general shadow = {
      shadow scale = 1,
      shadow xshift = -2ex,
      shadow yshift = 2ex,
      draw,
      thick,
      fill = white},
    general shadow = {
      shadow scale = 1,
      shadow xshift = -1ex,
      shadow yshift = 1ex,
      draw,
      thick,
      fill = white},
    fill = white, 
    draw,
    thick,
    minimum width = 2.5cm,
    minimum height = 1.3cm},
  squarednode/.style={rectangle, draw=green!60, fill=green!5, very thick, minimum width = 2.5cm, minimum height = 1.3cm}
}

\def\BeforeLight{2}
\begin{tikzpicture}
  % \draw[step=1cm, gray!20!white, very thin] (-5,-5) grid (5,5);

  % \node[draw, minimum height=1cm, minimum width=3cm] at (0, 3) (n_1) {Numerical Problem};
  \node[draw,ellipse,minimum width = 0.6cm, minimum height = 1.6cm, thick] at (0, 3) (problem) {\shortstack{Numerical\\Problem}};
  \node[draw,ellipse,minimum width = 0.6cm, minimum height = 1.6cm, thick] at (0, 0) (solution) {\shortstack{Raw Solution\\Data}};

  \node[cascaded] at (0.3,-3) (images) {Images};

  \node[draw,minimum width = 0.6cm, minimum height = 0.8cm, thick] at (2.0, 1.5) (solver) {\shortstack{CFD\slash PDE\\Solver}};
  \node[draw,minimum width = 0.6cm, minimum height = 0.8cm, thick] at (2.5, -1.5) (visu) {\shortstack{Visualization\\Tool}};

  \draw[->, shorten >=3pt, shorten <=3pt, thick ] (problem) edge (solution);
  \draw[->, shorten >=3pt, shorten <=3pt, thick ] (solution) edge (0, -2);

  \draw[-, dashed, shorten >=3pt, shorten <=3pt, thick ] (0.0, 1.5) edge (solver);
  \draw[-, dashed, shorten >=3pt, shorten <=3pt, thick ] (0.0, -1.5) edge (visu);

  \node[draw=green!60, fill=green!5, very thick, minimum width = 0.6cm, minimum height = 0.6cm] at (2, 4) (params_1) {\shortstack{Simulation\\Parameters}};
  \node[draw=green!60, fill=green!5, very thick, minimum width = 0.6cm, minimum height = 0.6cm] at (3.5, 2.3) (params_2) {\shortstack{Solver\\Parameters}};
  \node[draw=green!60, fill=green!5, very thick, minimum width = 0.6cm, minimum height = 0.6cm] at (3.5, -0.7) (params_3) {\shortstack{Visualization\\Parameters}};

\end{tikzpicture}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End: